Abstract

### 2025-10-09 output/Main_text.Rmd
Update "Graphical Representation Example" text
Make minor clarifications in the text.

Panels (a) and (b) show that,
although the agnostic coordinates are the same in both cases,
their relative position with respect to the latent trait distribution is not
---This is the effect of disregarding the covariance structure.
The correlation-based coordinates, plot in panels (c) and (d),
are also different in general from their agnostic counterpart (panel (b)).
This difference is especially salient in the location of items 2 and 3,
which are close to the 90% contour in the agnostic version,
but are almost along the 50% contour in the correlation-based one.
Panel (d) also illustrates the effect of plotting the coordinates directly
on the canonical basis;
the item locations and relative positions to the distribution
are correctly represented (allowing a more direct comparison with panel (b)),
but their angles with the axes and discrimination parameters are distorted:
For example, item 1 forms 45-degree angles with both axes,
instead of the 30-degree ones shown in Table \@ref(tab:example-items-table-out);
items 4 and 5 seem to be equally discriminating,
when Table \@ref(tab:example-items-table-out) panel (c) show they are not.

@@

We can also see how the sign of $`r INTERCEPT_PARAM`$ affects
the distance to the origin:
The item vector is applied at the origin when its value is null (item 1),
and tends to be shifted
*against* the item direction when its value is positive (items 2 and 4) and
*towards* the item direction when it is negative (items 3 and 5).
Finally, we can see how the direction is determined by the discrimination
parameters *and* the correlation; the sign of the discrimination parameters
determines where the item will point at;
thus, an item with two positive (negative) discrimination parameters will point
at the first (third) quadrant,
while an item with opposite-sign discrimination parameters will be in the
second or fourth quadrant.
The direction relative to the axes will be given by the relative absolute value
of the two parameters, but also by the correlation:
Items 2 and 3 especially illustrate this effect,
as each of them is orthogonal to one of the axes in the oblique space,
even when they are not parallel to the other axis.

>>>>

Panels (a) and (b) show that,
although the agnostic coordinates are the same in both cases,
their relative position with respect to the latent trait distribution is not
---this is the effect of disregarding the covariance structure.
The correlation-based coordinates, plot in panels (c) and (d),
are also different in general from their agnostic counterpart (panel (b)).
Although they do not vary in some cases (items 1 and 4),
the difference is especially salient in the location of items 2 and 3,
which are close to the 90% contour in the agnostic version
but almost on the 50% contour in the correlation-based one.
Panel (d) also illustrates the effect of plotting the coordinates directly
on the canonical basis;
the item locations and their positions relative to the distribution
are correctly represented (which makes comparing with panel (b) easier),
but their discrimination parameters and angles with the axes are distorted:
For example, items 4 and 5 seem to be equally discriminating,
when Table \@ref(tab:example-items-table-out) panel (c) show they are not;
and item 1 forms 45-degree angles with both axes,
instead of the 30-degree ones shown in Table \@ref(tab:example-items-table-out).

@@

We can also see how the sign of $`r INTERCEPT_PARAM`$ affects
the distance to the origin:
An item vector is applied at the origin when its value is null (item 1),
and tends to be shifted
*against* the item direction when its value is positive (items 2 and 4) and
*towards* the item direction when it is negative (items 3 and 5).
Finally, we can see how the direction is determined by the discrimination
parameters *and* the correlation; the sign of the discrimination parameters
determines where the item will point at;
thus, an item with two positive (negative) discrimination parameters will point
at the first (third) quadrant,
while an item with opposite-sign discrimination parameters will be in the
second or fourth quadrant.
The direction relative to the axes will be given by the relative absolute value
of the two parameters, but also by the correlation:
Items 2 and 3 especially illustrate this effect,
as each of them is orthogonal to one of the axes in the oblique space,
even when they are not parallel to the other one
